---
title: "Analyzing Course Data"
author: "DART team"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_float: yes
    fig_width: 5
    fig_height: 3
---

```{r load-packages, include=FALSE}
library(tidyverse)     #umbrella package which contains 
                       #ggplot for data visualization 
                       #and dplyr for data wrangling
```

## Introduction 

  We will examine course data which is readily available from our CANVAS gradebooks and also myUW. 
  

## The data 

Let's load the datasets 
```{r read-data}

gradebook <- read_csv("data/Grades_quarter1.csv", 
                      name_repair = "universal") %>%
             filter(is.na(Student) ) %>%
             select(-Student)

names(gradebook)

class_standing <- read_csv("data/myUW_quarter1.csv",
                           name_repair = "universal") %>% 
                  select(-Pronouns)

names(class_standing)
```

and join them by matching the UW NETID in `class_standing` with `SIS.Login.ID` in `gradebook`. 

```{r join-data}

course_data <- gradebook %>% 
       left_join(class_standing, by = c("SIS.Login.ID" = "UWNetID"))

glimpse(course_data)
```



1. How do exam points and non exam points correlate by class standing? 

```{r exam-versus-non-class}
ggplot(data = course_data,
       mapping = aes(x = TotalNonExamPercent,
                     y = TotalExamPercent,
                     color = Class)) +
  geom_point()

```

NOTES (08-15)
* Filter out the low data point
* add regression lines
* transform data using a logit or arcsine


2. How do exam points and non exam points correlate by major? 

```{r exam-versus-non-major}
# ggplot(data = course_data,
#        mapping = aes(x = TotalNonExamPercent,
#                      y = TotalExamPercent,
#                      color = Major)) +
#   geom_point()

```

## CFC NOTES:
* Group majors into buckets
* Show table or bar chart of numbers/percent of students in each year/major
* Box plots of exam/non-exam points by class standing/major
* Include (somewhere) explanations of the anatomy of the Rmd file
* Create metadata files for `class_standing` and `gradebook`. 